### 2014-07-30 paper/paper.tex
Made Meg's edits and upped the draft number

Utilizing Maximum Entropy classification for sentiment analysis in twitter
messages

@@

Draft 1 %TODO ditch this

@@

\newpage
\pagestyle{headings}
\setcounter{page}{1}

@@

Social Media allows people to discuss topics and disseminate their opinions
with a near negligible cost. Since the cost of using social media is so low it
allows everyone to share their opinions as much as they please. Because of this
the scale of textual data is enormous. In order to gain insight into the
`overall' opinion on particular topics, the process of determining sentiment
must be automated.

@@

Sentiment analysis is simply a classification task that attempts to label
fragments of text with a label that appropriately summarizes the attitude of
the text. Natural Language Processing (NLP) provides a number of method for
preforming sentiment analysis. Many of these methods, such as Naive Bayes and
Maximum Entropy, take a probabilistic approach to performing the task. Another
commonly used method of performing sentiment analysis is the uses of a support
vector machine (SVM).

@@

The focus of this undergraduate research project has been to preform sentiment
analysis in data from the microbloging website Twitter. This data presents some
interesting problems to performing good sentiment analysis due to it's short
and informal nature. All of the messages, known as Tweets, are constrained in
length to 140 characters. This limit forces many users of Twitter to use
abbreviations and contractions. Twitter makes it extremely easy to publish a
Tweet and has no way to edit a Tweet once published. This results more spelling
and grammar mistakes then other sources of text. Tweets often also contain
links to other websites in the form of URLs as well as taging system for topics
and other users.

@@

Organizers of the SemEval-2013 workshop created various sentiment analysis
tasks to be preformed. The task focused on for this project was task-2b:
Message polarity classification. The organizers used the amazon service
``Mechanical Turk" to get many humans to annotate Tweets as `positive',
`negative' or `neutral'. In tweets with mixed sentiments the prevailing
sentiment was chosen. Participants were tasked with creating a classifier to
establish labels for Tweets not in the training dataset. The various
submissions of the participants were then evaluated by the average of the
F-positive and the F-negative. The submissions were allowed categorized as
`unconstrained' or `constrained' depending on if they did or did not use
supplemental data, respectively.

@@

@@

The Maximum Entropy classification method that was focused on due to the
limited duration of the project. This supervised machine learning method
calculates the probability that a given document $d$ belongs to the class $c$.
The maximum entropy approach is well described in \cite{Nigam1999} and is
summarized below.

The training data is used to establish constraints in the following fashion.

@@

Where $Z(d)$ is a normalization function, $F_{i,c}$ is a feature function for
feature $f_{i}$ and class $c$, and $\lambda_{i}$ is a feature parameter.

@@

The classifier produced for this project is implemented in \textit{python} and
make extensive use of the \textit{nltk} (Natural Language Toolkit) library. As
well the implementation makes use of the \textit{flex} tool to create a lexer
for tokenization. The classifier does not make uses of any supplemental
datasets and therefore should be considered constrained within the context of
the task.

@@

    \item \textbf{Splitting} - The corpus of documents are random separated
        into three sets: the \textit{training set}, the \textit{validation set},
        and the \textit{testing set}. Respectively, Each dataset is 60\%, 20\%,
        and 20\% of the corpus.

@@

The implementation created for this project used a validation set during
training as the primary mechanism to avoid overfitting. While training the
effectiveness of the classifier was evaluated after every iteration, using the
accuracy or F-positive measure. During the training process the following
procedure was followed:

@@

After the SemEval-2013 workshop the organizers released a paper discussing the
proceedings and the results of the task \cite{Nakov2013}. The paper establishes
baseline in terms of $F$. $F$ can be calculated $F = (F_{pos} + F_{neg}) / 2$
where $F_{pos} = 2 \frac{P_{pos}R_{pos}}{P_{pos} + R_{pos}}$ and $F_{neg}$ is
calculated in a similar fashion. The Precision measure for the positive label
is $P_{pos} = \frac{\mathrm{\# of\ true\ postive}}{\mathrm{\# of\ true\
postive} + \mathrm{\# of\ false\ postive} }$. The Recall measure for the
positive label is defined as $R_{pos} = \frac{\mathrm{\# of\ true\
postive}}{\mathrm{\# of\ true\ postive} + \mathrm{\# of\ false\ negative} }$

@@

As well, some Twitter users may be frequently be referred to with a particular
sentiment. For a hypothetical example, Tweets addresses to Emma Watson
(`@EmWatson') are more likely to have a positive sentiment, while tweets to
Barack Obama (`@BarackObama') may contain extremely mixed sentiment.

>>>>

Utilizing Maximum Entropy Classification for Sentiment Analysis in Twitter
Messages

@@

Draft 2 %TODO ditch this

@@

\newpage
%\pagestyle{headings}
\setcounter{page}{1}

@@

Social media allows people to discuss topics and disseminate their opinions
with a near negligible cost. The low barrier of entry to social media allows
everyone to share their opinions as much as they please. The scale of textual
data from social media is enormous. In order to gain insight into the `overall'
opinion on particular topics, the process of determining sentiment must be
automated.

@@

Sentiment analysis is simply a classification task that attempts to label
fragments of text that appropriately summarizes the attitude of the text.
Natural Language Processing (NLP) provides a number of method for preforming
sentiment analysis. Many of these methods, such as Naive Bayes and Maximum
Entropy, take a probabilistic approach to performing the task. Another commonly
used method of performing sentiment analysis is the uses of a support vector
machine (SVM).

@@

The focus of this undergraduate research project has been to preform sentiment
analysis in data from the microbloging website Twitter. All of the messages,
known as Tweets, are constrained in length to 140 characters. This limit forces
many users of Twitter to use abbreviations and contractions. This data presents
some interesting problems to performing good sentiment analysis due to it's
short and informal nature. Twitter makes it extremely easy to publish a Tweet
and has no way to edit a Tweet once published. This results in more spelling
mistakes and inconsistent grammar then other sources of text. Tweets often also
contain links to other websites in the form of URLs as well as taging system
for topics and other users.

@@

Organizers of the SemEval-2013 workshop created various sentiment analysis
tasks to be preformed. The task focused on for this project was task-2b:
Message polarity classification. The organizers used the Amazon service
``Mechanical Turk" to get many humans to annotate Tweets as `positive',
`negative' or `neutral'. In tweets with mixed sentiments, the prevailing
sentiment was chosen. Participants were tasked with creating a classifier to
establish labels for Tweets not in the training dataset. The various
submissions of the participants were then evaluated by the average of the
F-positive and the F-negative. The submissions were allowed categorized as
`unconstrained' or `constrained' depending on if they did or did not use
supplemental data, respectively.

@@

\textit{Note: This section needs to be expanded. I am just not sure what else to write about...}

@@

The Maximum Entropy classification method was focused on due to the limited
duration of the project. This supervised machine learning method calculates the
probability that a given document $d$ belongs to the class $c$.  The Maximum
Entropy approach is summarized below. For a more detailed description, see:
Nigam et al. \cite{Nigam1999}.

The training data is used to establish constraints in the following fashion:

@@

where $Z(d)$ is a normalization function, $F_{i,c}$ is a feature function for
feature $f_{i}$ and class $c$, and $\lambda_{i}$ is a feature parameter.

@@

The classifier produced for this project is implemented in \textit{python} and
make extensive use of the \textit{nltk} (Natural Language Toolkit) library. As
well the implementation makes use of the \textit{flex} tool to create a lexer
for tokenization. The classifier does not make uses of any supplemental
datasets and therefore should be considered constrained within the context of
the task. The classifier adheres to the following procedure:

@@

    \item \textbf{Splitting} - The corpus of documents are random separated
        into three sets: the \textit{training set}, the \textit{validation set},
        and the \textit{testing set}. Respectively, each dataset is 60\%, 20\%,
        and 20\% of the corpus.

@@

The implementation created for this project used a \textit{validation set} during
training as the primary mechanism to avoid overfitting. While training the
effectiveness of the classifier was evaluated after every iteration, using the
accuracy or F-positive measure. During the training process the following
procedure was followed:

@@

After the SemEval-2013 workshop the organizers released a paper discussing the
proceedings and the results of the task \cite{Nakov2013}. The paper establishes
baseline in terms of $F$. $F$ can be calculated $F = (F_{pos} + F_{neg}) / 2$
where $F_{pos} = 2 \frac{P_{pos}R_{pos}}{P_{pos} + R_{pos}}$ and $F_{neg}$ is
calculated in a similar fashion. The \textit{Precision} measure for the
positive label is $P_{pos} = \frac{\mathrm{\# of\ true\ postive}}{\mathrm{\#
of\ true\ postive} + \mathrm{\# of\ false\ postive} }$. The \textit{Recall}
measure for the positive label is defined as $R_{pos} = \frac{\mathrm{\# of\
true\ postive}}{\mathrm{\# of\ true\ postive} + \mathrm{\# of\ false\ negative}
}$

@@

As well, some Twitter users may be frequently be referred to with a particular
sentiment. For a hypothetical example, Tweets addressed to Emma Watson
(`@EmWatson') are more likely to have a positive sentiment, while tweets to
Barack Obama (`@BarackObama') may contain extremely mixed sentiment.
